Food-web structure, stochasticity and response diversity drive temporal community stability in complex communities

### 2024-02-16 _targets.Rmd
Update figures

```{targets rho_model}
list(
  tar_target(formula_fixed_rho, "log_richness +
    ct_alive +
    log1_avg_max_int_alive +
    productivity +
    rho +
    I(rho^2) +
    env_stoch +

    log_richness:ct_alive +
    log_richness:log1_avg_max_int_alive +
    log_richness:productivity +
    log_richness:rho +

    log1_avg_max_int_alive:ct_alive +
    log1_avg_max_int_alive:productivity +
    log1_avg_max_int_alive:rho +

    ct_alive:rho +
    ct_alive:productivity +
    rho:productivity +
    I(rho^2):productivity +

    log_richness:log1_avg_max_int_alive:ct_alive +
    log_richness:log1_avg_max_int_alive:productivity +
    log_richness:log1_avg_max_int_alive:rho +
    log_richness:ct_alive:productivity +
    log_richness:ct_alive:rho +
    log_richness:rho:productivity +

    log1_avg_max_int_alive:ct_alive:rho +
    log1_avg_max_int_alive:ct_alive:productivity +
    log1_avg_max_int_alive:rho:productivity +

    ct_alive:rho:productivity"
    ),
  tar_target(formula_rand_rho, "(1 | Z)"),
  tar_target(mod_rho, map(setNames(
    c("log_stab_com", "log_stab_pop", "log_async"),
    c("log_stab_com", "log_stab_pop", "log_async")
    ),
  function (x) {
    glmmTMB(formula(paste0(x, " ~ ", formula_fixed_rho, " + ", formula_rand_rho)),
      sim_rho)
  }
  )),
  tar_target(coef_mod_rho, map_dfr(mod_rho, ~model2tibble(.x))),
  tar_target(r2_mod_rho, map_dfr(mod_rho, ~as.data.frame(r2_nakagawa(.x)), .id = "response") %>%
  mutate(response = var_replacement()[response])),
  tar_target(p_mod_rho, plot_effect_type(coef_mod_rho)),
  tar_target(mod_rho_scaled, map(setNames(
    c("log_stab_com", "log_stab_pop", "log_async"),
    c("log_stab_com", "log_stab_pop", "log_async")
    ),
  function (x) {
    glmmTMB(formula(paste0(x, " ~ ", formula_fixed_rho, " + ", formula_rand_rho)),
      sim_rho_scaled)
  }
)
    ),
  tar_target(coef_mod_rho_scaled, map_dfr(mod_rho_scaled, ~model2tibble(.x))),
  tar_target(r2_mod_rho_scaled,
    map_dfr(mod_rho_scaled, ~as.data.frame(r2_nakagawa(.x)), .id = "response") %>%
      mutate(response = var_replacement()[response])),
  tar_target(colin_rho, performance::check_collinearity(lm(formula(str_replace_all(paste0("log_stab_com ~ ", formula_fixed_rho), "(\\*|:)", "+")),
  sim_rho_scaled))),
    tar_target(p_mod_rho_scaled, plot_effect_type(coef_mod_rho_scaled))
)
```

@@

@@

```{targets p-stab-rich}
list(
#  tar_target(p_stab_s_prod,
#    sim_fw %>%
#      filter(productivity %in% c(5, 30), rho == 0, env_stoch == .3) %>%
#      ggplot(aes(x = richness, y = stab_com, color = as.factor(productivity))) +
#      geom_point() +
#      geom_smooth() +
#      labs(x = "Species richness", y = "Community stability", color = "Productivity") +
#      theme_bw()),
    tar_target(p_stab_s_env_stoch,
      sim_fw %>%
        filter(env_stoch %in% c(.5, 1.5), rho == 0) %>%
        ggplot(aes(x = richness, y = stab_com, color = as.factor(env_stoch))) +
        geom_point() +
        geom_smooth() +
        labs(x = "Species richness", y = "Community stability", color = "Env. variability") +
        theme_bw()),
      tar_target(p_stab_s_resp_div,
        sim_fw %>%
          filter(rho %in% c(0, 1), env_stoch == .5) %>%
          ggplot(aes(x = richness, y = stab_com, color = as.factor(1 - rho))) +
          geom_point() +
          geom_smooth() +
          labs(x = "Species richness", y = "Community stability", color = "Response diversity") +
          theme_bw()
      ),
      tar_target(p_stab_s_stress,
         plot_grid(
#           p_stab_s_prod,
           p_stab_s_env_stoch,
           p_stab_s_resp_div,
           labels = "auto"
         )
      )
)
```

@@

```{targets p_sem_tot_stab, tar_simple=TRUE}
tar_read(sem_eff_tab) %>%
  filter(effect_type == "total") %>%
  filter(response %in% c("async", "pop_stab", "stab_com")) %>%
  filter(!predictor %in% c("evenness_sae", "sae_even", "cpe_int", "cpe_env", "pop_stab", "async", "S", "ct")) %>%
  mutate(across(c(response, predictor), ~var_replacement()[.x])) %>%
  ggplot(aes(
      x = predictor,
      y = effect,
      ymin = lower_ci,
      ymax = upper_ci)
    ) +
  geom_pointrange() +
  geom_hline(yintercept = 0, linetype = "dashed") +
  facet_grid(cols = vars(response)) +
  labs(x = "Predictor", y = "Total effect") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 20, hjust = 1))
```

@@

```{r tar-make}
#tar_make()
tar_make_future(workers = 3)
```

@@

```{r}
tar_load(c(formula_sem_bivariate_relations, scaled_sem_simplified, scaled_sim_fw_sem))
sem_simplified_bivariate_relations <- tibble(formula = formula_sem_bivariate_relations) %>%
      separate(
        col = formula,
        into = c("response", "predictor"),
        sep = " ~ ", remove = FALSE) %>%
    mutate(
      data = map(formula_sem_bivariate_relations,
        ~partialResid(as.formula(.x),
            modelList = scaled_sem_simplified,
            data = scaled_sim_fw_sem
          )
        ),
      coefs = map2(response, predictor,
        ~coefficients(scaled_sem_simplified[[.x]])[c("(Intercept)", .y)]
        ),
      plot = pmap(list(
          data = data,
          response = response,
          predictor = predictor,
          coef = coefs
          ),
        function (data, response, predictor, coef, ext_data) {
          if (predictor %in% c("h", "env_stoch", "resp_div")) {
            p <- data %>%
              ggplot(aes(
                  y = yresid, x = xresid,
                  group = cut(round(xresid, 1),
                    breaks = length(unique(ext_data[[predictor]]))
                  )
                  )) +
              geom_boxplot(alpha = .1)
          } else {
            p <- data %>%
              ggplot(aes(y = yresid, x = xresid)) +
              geom_point(alpha = .1)
          }
          p +
            geom_abline(
              intercept = coef["(Intercept)"],
              slope = coef[2],
              linewidth = 1
              ) +
            labs(
              x = var_replacement()[predictor],
              y = var_replacement()[response]
              ) +
            coord_cartesian(expand = FALSE) +
            theme_half_open()

@@

```{r}
label_sem <- as.character(as.roman(seq_len(nrow(sem_simplified_bivariate_relations))))
p_sem_simplified_bivariate_relations <-
  plot_grid(
    plotlist = sem_simplified_bivariate_relations$plot,
    label_x = 1, hjust = 1.0,
    labels = label_sem
  )
sem_file <- here("report", "figures", "sem_stab_simplified.pdf")
p_sem <- ggdraw() +
  draw_image(magick::image_read_pdf(sem_file, density = 600),
    x = 0, y = 0,
    width = 1, height = 1,
    hjust = 0, halign = 0, valign = 0
  )
p_sem_tot <- plot_grid(
  p_sem, p_sem_simplified_bivariate_relations,
  nrow = 2, rel_heights = c(.6, 1),
  labels = "auto"
)
ggsave_multiple(
  paste0("p_sem_tot", c(".jpg", ".png", ".pdf")),
  plot = p_sem_tot,
  path = here::here("report", "figures"),
  scale = 2.4,
  dpi = 400,
  width = 80,
  height = 80 * 1.6,
  units = "mm")
```

@@

```{r}
library(RColorBrewer)
library(ggbreak)
#cols <- c("Community stability" = "black",
#  "Asynchrony" = "blue",
#  "Population stability" = "darkgreen"
#)
custom_stab_color <- c(
  "Community stability" = "black", #"#ffffbf"
  "Asynchrony" = "#91bfdb",
  "Population stability" = "#fc8d59"
)
alpha <- c("Community stability" = 1, "Asynchrony" = 1, "Population stability" = 1)

sem_tot_predictor <- c("richness", "avg_int_strength", "w_avg_tlvl",
  "ct_alive", "Z", "h", "resp_div", "env_stoch")
var_replacement()[sem_tot_predictor]
p_sem_tot_stab <- tar_read(sem_simplified_eff_tab) %>%
  filter(effect_type == "total") %>%
  filter(response %in% c("async", "pop_stab", "stab_com")) %>%
  filter(predictor %in% sem_tot_predictor) %>%
  mutate(
    response = recode_factor(response, !!!var_replacement(), .ordered = TRUE),
    predictor = recode_factor(predictor, !!!var_replacement()[rev(sem_tot_predictor)], .ordered = TRUE)
) %>%
  ggplot(aes(
      y = predictor,
      x = effect,
      color = response,
      alpha = response,
      ymin = lower_ci,
      ymax = upper_ci)
    ) +
  geom_pointrange() +
  scale_colour_manual(values = custom_stab_color) +
  scale_alpha_manual(values = alpha) +
#  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_vline(xintercept = 0, linetype = "dashed") +
  #facet_grid(cols = vars(response)) +
  labs(
    x = "Total standardised effects",
    color = "Stability component",
    alpha = "Stability component"
    ) +
  theme_bw() +
  theme(
    axis.title.y = element_blank(),
    #legend.justification = "right",
    legend.position = "bottom") +
  guides(
    color = guide_legend(override.aes = list(linetype = 0, alpha = 1))
    ) +
  theme(
    legend.title = element_blank(),
    legend.position = c(.005, .995),
    legend.justification = c("left", "top")
  )
p_sem_tot_stab_leg <- plot_grid(
  print(p_sem_tot_stab +
    theme(
      legend.position = "none",
    )
  ),
  get_legend(p_sem_tot_stab),
  nrow = 2,
  rel_heights = c(1, .05)
)
ggsave_multiple(
  paste0("p_sem_tot_stab", c(".png", ".pdf")),
  plot = print(
    p_sem_tot_stab +
      guides(
        color = guide_legend(override.aes = list(linetype = 0, alpha = 1))
      ) +
      theme(
        legend.title = element_blank(),
        legend.position = c(.005, .995),
        legend.justification = c("left", "top"),

@@

p_sem_fig <- plot_grid(
  p_sem, p_bivariate_rel,
  print(
    p_sem_tot_stab +
      guides(
        color = guide_legend(override.aes = list(linetype = 0, alpha = 1))
      ) +
      theme(
        legend.title = element_blank(),
        legend.position = c(.005, .995),
        legend.justification = c("left", "top"),

@@

p_sem_fig <- plot_grid(p_sem,
  print(
    p_sem_tot_stab +
      guides(
        color = guide_legend(override.aes = list(linetype = 0, alpha = 1))
      ) +
      theme(
        legend.title = element_blank(),
        legend.position = c(.005, .995),
        legend.justification = c("left", "top"),

>>>>

```{targets rho_model}
list(
  tar_target(formula_fixed_rho, formula_mod_rho_fixed_effect()),
  tar_target(formula_rand_rho, "(1 | Z)"),
  tar_target(mod_rho, map(setNames(
    c("log_stab_com", "log_stab_pop", "log_async"),
    c("log_stab_com", "log_stab_pop", "log_async")
    ),
  function (x) {
    glmmTMB(formula(paste0(x, " ~ ", formula_fixed_rho, " + ", formula_rand_rho)),
      sim_rho)
  }
  )),
  tar_target(coef_mod_rho, map_dfr(mod_rho, ~model2tibble(.x))),
  tar_target(r2_mod_rho, map_dfr(mod_rho, ~as.data.frame(r2_nakagawa(.x)), .id = "response") %>%
  mutate(response = var_replacement()[response])),
  tar_target(p_mod_rho, plot_effect_type(coef_mod_rho)),
  tar_target(mod_rho_scaled, map(setNames(
    c("log_stab_com", "log_stab_pop", "log_async"),
    c("log_stab_com", "log_stab_pop", "log_async")
    ),
  function (x) {
    glmmTMB(formula(paste0(x, " ~ ", formula_fixed_rho, " + ", formula_rand_rho)),
      sim_rho_scaled)
  }
)
    ),
  tar_target(coef_mod_rho_scaled, map_dfr(mod_rho_scaled, ~model2tibble(.x))),
  tar_target(r2_mod_rho_scaled,
    map_dfr(mod_rho_scaled, ~as.data.frame(r2_nakagawa(.x)), .id = "response") %>%
      mutate(response = var_replacement()[response])),
  tar_target(colin_rho, performance::check_collinearity(lm(formula(str_replace_all(paste0("log_stab_com ~ ", formula_fixed_rho), "(\\*|:)", "+")),
  sim_rho_scaled))),
    tar_target(p_mod_rho_scaled, plot_effect_type(coef_mod_rho_scaled))
)
```

@@

```{targets sem_tot_predictor, tar_simple = TRUE}
c("richness", "avg_int_strength", "w_avg_tlvl", "ct_alive", "Z", "h",
  "resp_div", "env_stoch")
```

```{targets plot-sem-tot}
list(
  tar_target(p_sem_tot_stab,
    sem_simplified_eff_tab %>%
      filter(effect_type == "total") %>%
      filter(response %in% c("async", "pop_stab", "stab_com")) %>%
      filter(predictor %in% sem_tot_predictor()) %>%
      mutate(
        response = recode_factor(response, !!!var_replacement(), .ordered = TRUE),
        predictor = recode_factor(predictor, !!!var_replacement()[rev(sem_tot_predictor())], .ordered = TRUE)
        ) %>%
      ggplot(aes(
          y = predictor,
          x = effect,
          color = response,
          alpha = response,
          ymin = lower_ci,
          ymax = upper_ci)
        ) +
      geom_pointrange() +
      scale_colour_manual(values = stab_color()) +
      scale_alpha_manual(values = stab_alpha()) +
      #  geom_hline(yintercept = 0, linetype = "dashed") +
      geom_vline(xintercept = 0, linetype = "dashed") +
      #facet_grid(cols = vars(response)) +
      labs(
        x = "Total standardised effects",
        color = "Stability component",
        alpha = "Stability component"
        ) +
      theme_bw() +
      theme(
        axis.title.y = element_blank(),
        #legend.justification = "right",
        legend.position = "bottom") +
      guides(
        color = guide_legend(override.aes = list(linetype = 0, alpha = 1))
        ) +
      theme(
        legend.title = element_blank(),
        legend.position = c(.005, .995),
        legend.justification = c("left", "top")
      )
    ),
  tar_target(p_sem_tot_stab_leg,
    plot_grid(
      print(p_sem_tot_stab +
        theme(
          legend.position = "none",
        )
        ),
      get_legend(p_sem_tot_stab),
      nrow = 2,
      rel_heights = c(1, .05)
      )
  )
)
```

@@

```{targets p-stab-rich}
list(
    tar_target(p_stab_s_env_stoch,
      sim_fw %>%
        filter(env_stoch %in% c(.5, 1.5), rho == 0) %>%
        ggplot(aes(x = richness, y = stab_com, color = as.factor(env_stoch))) +
        geom_point() +
        geom_smooth() +
        labs(x = "Species richness", y = "Community stability", color = "Env. variability") +
        theme_bw()),
      tar_target(p_stab_s_resp_div,
        sim_fw %>%
          filter(rho %in% c(0, 1), env_stoch == .5) %>%
          ggplot(aes(x = richness, y = stab_com, color = as.factor(1 - rho))) +
          geom_point() +
          geom_smooth() +
          labs(x = "Species richness", y = "Community stability", color = "Response diversity") +
          theme_bw()
      ),
      tar_target(p_stab_s_stress,
         plot_grid(
           p_stab_s_env_stoch,
           p_stab_s_resp_div,
           labels = "auto"
         )
      )
)
```

@@

@@

```{r tar-make}
tar_make()
tar_visnetwork()
tar_make_future(workers = 3)
```

@@

## SEM plot

The partial residual function does not work inside targets unfortunately,

>   Last error: ℹ In argument: `data = map(...)`.
> Caused by error in `map()`:
> ℹ In index: 1.
> Caused by error in `is.data.frame()`:
> ! object 'scaled_sim_fw_sem' not found

```{r}
tar_load(
  c(
    formula_sem_bivariate_relations,
    scaled_sem_simplified,
    scaled_sim_fw_sem
  )
)
sem_simplified_bivariate_relations <- tibble(formula = formula_sem_bivariate_relations) %>%
  separate(
    col = formula,
    into = c("response", "predictor"),
    sep = " ~ ", remove = FALSE
    ) %>%
mutate(
  data = map(formula_sem_bivariate_relations,
    ~partialResid(as.formula(.x),
      modelList = scaled_sem_simplified,
      data = scaled_sim_fw_sem
    )
    ),
  coefs = map2(response, predictor,
    ~coefficients(scaled_sem_simplified[[.x]])[c("(Intercept)", .y)]
    ),
  plot = pmap(list(
      data = data,
      response = response,
      predictor = predictor,
      coef = coefs
      ),
    function (data, response, predictor, coef, ext_data) {
      if (predictor %in% c("h", "env_stoch", "resp_div")) {
        p <- data %>%
          ggplot(aes(
              y = yresid, x = xresid,
              group = cut(round(xresid, 1),
                breaks = length(unique(ext_data[[predictor]]))
              )
              )) +
          geom_boxplot(alpha = .1)
      } else {
        p <- data %>%
          ggplot(aes(y = yresid, x = xresid)) +
          geom_point(alpha = .1)
      }
      p +
        geom_abline(
          intercept = coef["(Intercept)"],
          slope = coef[2],
          linewidth = 1
          ) +
        labs(
          x = var_replacement()[predictor],
          y = var_replacement()[response]
          ) +
        coord_cartesian(expand = FALSE) +
        theme_half_open()

@@

```{r}
label_sem <- as.character(as.roman(seq_len(nrow(sem_simplified_bivariate_relations))))
p_sem_simplified_bivariate_relations <- plot_grid(
  plotlist = sem_simplified_bivariate_relations$plot,
  label_x = 1, hjust = 1.0,
  labels = label_sem
)
```

```{r}
sem_file <- here("report", "figures", "sem_stab_simplified.pdf")
p_sem <- ggdraw() +
  draw_image(magick::image_read_pdf(sem_file, density = 600),
    x = 0, y = 0,
    width = 1, height = 1,
    hjust = 0, halign = 0, valign = 0
  )
p_sem_tot <- plot_grid(
  p_sem, p_sem_simplified_bivariate_relations,
  nrow = 2, rel_heights = c(.6, 1),
  labels = "auto"
)
```

```{r fig-sem}
ggsave_multiple(
  paste0("p_sem_tot", c(".jpg", ".png", ".pdf")),
  plot = p_sem_tot,
  path = here::here("report", "figures"),
  scale = 2.4,
  dpi = 400,
  width = 80,
  height = 80 * 1.6,
  units = "mm")
```

@@

```{r}
ggsave_multiple(
  paste0("p_sem_tot_stab", c(".png", ".pdf")),
  plot = print(
    tar_read(p_sem_tot_stab) +
      guides(
        color = guide_legend(override.aes = list(linetype = 0, alpha = 1))
      ) +
      theme(
        legend.title = element_blank(),
        legend.position = c(.005, .995),
        legend.justification = c("left", "top"),

@@

p_sem_fig <- plot_grid(
  p_sem, p_bivariate_rel,
  print(
    tar_read(p_sem_tot_stab) +
      guides(
        color = guide_legend(override.aes = list(linetype = 0, alpha = 1))
      ) +
      theme(
        legend.title = element_blank(),
        legend.position = c(.005, .995),
        legend.justification = c("left", "top"),

@@

p_sem_fig <- plot_grid(p_sem,
  print(
    tar_read(p_sem_tot_stab) +
      guides(
        color = guide_legend(override.aes = list(linetype = 0, alpha = 1))
      ) +
      theme(
        legend.title = element_blank(),
        legend.position = c(.005, .995),
        legend.justification = c("left", "top"),

### 2024-02-16 paper/manuscript.Rmd
Update figures

---
## IDEA, meeting with Sonia
#- Simplifying SEM in main text, move details to appendix
#- State diagram for stability - richness relationships along two food-web
#  metrics and panels varying response diversity.
#- Look at the french data for portefolio and compensatory dynamics effects.
#- Mid-October
---

>>>>

---
## IDEA:
# Talk about about how stochasticity minor interaction effects similar to
# dispersal (Holt, 1985; other papers I guess)
---
